Learning with misspecified models: the case of overestimation

### 2023-10-14 Paper/draft.Rmd
smm

---
title: "Learning with misspecified models: Overconfidence and Stereotypes"  # nolint: Grammarly.
date: '`r format(Sys.time(), "%B %d, %Y")`'
author: "Jimena Galindo"
abstract: "TBW"
#classoption: pagebackref # passes pagebackref=true to hyperref before it is loaded, allowing backreferencing
output: 
    pdf_document:
        citation_package: natbib
        fig_width: 7  
        fig_height: 6
        fig_caption: true
        number_sections: true

        template: NULL
        keep_tex: true
        extra_dependencies:
            footmisc: ["bottom"] # footnote management 
            setspace: ["doublespacing"] # spacing of the paper 
            caption: ["normal"] 
            dsfont: null # indicator function 1
            booktabs: null 
            makecell: null 
            hyperref: null 
        includes:
          in_header: extra_header.tex

bibliography: references.bib
fontsize: 12pt 
geometry: margin=2.5cm
---

@@

# Introduction

Overconfidence has been documented in a wide range of settings and has been shown to have important economic consequences:
@Camerer1999 find that overconfidence can explain the excessive entry of entrepreneurs into the market; @Hoffman2020 show that
workers who overestimate their productivity within a firm are less likely to look for a new job; @Oster2013 determine that 
people at high risk of Huntington's disease make saving decisions as if they were not at risk.

Although the bias has been widely studied by economists, the settings in which 
the the belief updating procedure has been studied is narrow in scope. In particular, the literature has focused on
the cases in which the agent updates their beliefs about a single characteristic that they might care about.
For instance, @Eil2011 study the case in which the agent updates their beliefs about their own ability, and their 
own beauty in two separate experiments. Similarly, @Benoit2015 study the case in which the agent updates their 
beliefs about their own IQ.

@@

I consider multiple theories of belief updating that have been proposed in the literature to rationalize
the prevalence of misspecified beliefs. I focus on two different frameworks and develop a unifying example that allows me to
compare the predictions of all the theories. In framework 1 I consider a setting in which the an agent observes
a continuous output and infers the underlying state. In framework 2, the agent observes a binary outcome and infers the
underlying probability of success. In both frameworks, I consider two nested theories of belief updating.

@@

_The student must choose an effort level $a$, which determines her grade. For whatever the chosen effort level, 
the student must pay a cost $c(a) = \frac{1}{2}a^2$ and receives a grade of $(\theta^*+a)\omega^*$. She repeats 
this process for infinitely many periods. 
Assume also that the student's prior is such that $\mathbb{E}[\omega]= \omega^*$ and she is dogmatic 
about being of 
type $\hat{\theta}>\theta^*$.\footnote{The example is illustrated for an overconfident student but 
the results are symmetric for a dogmatic student who initially 
places probability one on some $\tilde{\theta}<\theta^*$.}
Therefore, the student's payoff in period $t$ is given by _

@@

_Under this specification, the myopic optimal effort level is $a_t^* = \omega^*$. 
Because the student does not know $\omega^*$,
he will choose $a_t = \mathbb{E}_t(\omega)$ where the expectation is taken with respect to the 
student's belief at the beginning of
period $t$. If she does not revise his effort choice for $k$ periods, she will receive 
an average utility of 
$(\theta^*+a_t^*)\omega^* - \frac{1}{2}a_t^{*2}$ but he was expecting an average utility of 
$(\hat{\theta}\theta+a_t^*)\omega^* - \frac{1}{2}a_t^{*2}$. In response, she will apply Bayes 
rule to update his beliefs about
$\omega$ to get the posterior belief with 
$\mathbb{E}_{t+k}[\omega] = \frac{(\theta^{*} + \omega^{*})\omega^{*}}{\hat{\theta} + \omega^{*}}$
which is lower than the initial belief. This will cause the student to choose a lower effort at $t+k$. As a result, she will again
receive an average utility that is lower than what he expected which will cause her belief to drift further down. This process will
continue until the average utility equals her expected utility under the dogmatic belief that assigns probability 1 to $\hat{\theta}$. 
At that point, the student will have reached a self-defeating equilibrium and she will continue to choose sub-optimal effort forever. _

@@

The overall pattern is that bad news are associated with a decrease in effort, and a 
more negative news differential is associated with a larger decrease in effort. On the contrary, good news are associated with 
a small decrease in effort. As mentioned in the predictions section, this is consistent with the self-attribution bias as well
as with a change of paradigm. It is not consistent with the dogmatic modeler or the Bayesian. Because 
I do not directly elicit beliefs, it is not possible to distinguish between the self-attribution bias and the change of 
paradigm with this data alone.
Nevertheless, the fact that about a third of the subjects started with a correctly specified belief and ended up overconfident
is consistent only with the self-attribution bias. Therefore, I conclude that the self-attribution bias is the most likely
explanation for the behavior of the subjects in the experiment.

@@

A regression analysis of this same data reveals that indeed, the slopes are different for the two types 
of news and that the difference is statistically significant ($p<0.01$). The regression results are reported in Table \ref{tab:regression}.
columns 2 and 3 of the table consider the model for each treatment separately. I find no significant differences in 
the parameters across treatment. This is consistent with the fact that the bias is not driven by the ego-relevance 
of $\theta$.\footnote{The ego-relevance of $\theta$ is the motivation behind the self-attribution bias, 
however, the bias itself is due to an incorrect updating procedure in response to good and bad news. The presence of 
asymmetric updating in the stereotype condition can be due to self-censoring of the subjects. In particular, subjects 
who initially underestimated the performance of the other participant might have realized their mistake and 
by overestimating it after good news. This is consistent with the data presented here as well as with 
the data observed in the transition matrix for stereotypes which is reported in the Appendix}

@@

More convincing evidence against the Switching model is the lack of inverse or attenuated correlation between 
the change in effort and the news differential. In particular, if the switcher were the underlying model,
I would expect overconfident subjects who start with an overestimation of $\theta$ to increase their effort
in response to bad news that were very surprising. This would flatten the slope of the regression line 
for bad news relative to what a Bayesian would do. \ref{tab:regression} shows that the slope for bad news for
the simulation\footnote{The details on how the data was simulated are explained in the Appendix} 
of the Bayesian updates is much flatter than the slope in the data.

@@

@@

## Model fit

@@

I have small or no effects throughout on the ego-relevance of the type. Although the motivation behind the 
theories does not directly apply to the case of stereotypes, the lack of stark differences suggest that 
biases arise due to an incorrect updating procedure and not necessarily due to the self-serving motivation 
behind the bias. The lack of differences in the behavior of the subjects across treatments could also 
be due to self-censoring of the subjects. In particular, subjects might be initially reluctant to 
express their true belief about a particular group of people. In addition, as they learn and realize 
that their beliefs might be incorrect, they might overreact to signals to try to compensate for the initial 
bias. Such an updating procedure would be consistent with the self-attribution bias for the case of sterotypes.
I remain agnostic about whether that is truly the case or whether the lack of differences is due to other 
factors.

@@

# Appendix

## Transition matrix by treatment

@@

@@

# Experiment Instructions

@@

_the full data generating process is printed in the rear side of the instructions and always available to the subjects the matrices are labeled with "if your/the other participant's score is" depending on the treatment_

@@

# References

>>>>

---
title: "Learning with misspecified models: the case of overestimation"  
date: '`r format(Sys.time(), "%B %d, %Y")`'
author: "Jimena Galindo"
abstract: "I design a framework and laboratory experiment that allows for the comparison of multiple theories of misspecified learning. I focus on a framework with endogenous information and two a data generating process ruled by two fundamentals: an ego-relevant parameter and a state. Within this framework I study three forces that can lead to misspecified beliefs: initial misspecifications, learning traps and biased updating. I find that biased updating is the main driver of misspecifications. In addition, I vary the degree of ego-relevance of the parameter by introducing a stereotype treatment. The data is consistent with biased updating in both cases but for potentially different reasons: In the ego-relevant treatment subjects attribute successes to their own ability and failures to luck; in the stereotype treatment they over-compensate for initial negative biases by over-attributing positive signals to the ability of others. This translates into similar observed choices but different dynamic in beliefs."
#classoption: pagebackref # passes pagebackref=true to hyperref before it is loaded, allowing backreferencing
output: 
    pdf_document:
        citation_package: natbib
        fig_width: 7  
        fig_height: 6
        fig_caption: true
        number_sections: true

        template: NULL
        keep_tex: true
        extra_dependencies:
            footmisc: ["bottom"] # footnote management 
            setspace: ["doublespacing"] # spacing of the paper 
            caption: ["normal"] 
            dsfont: null # indicator function 1
            booktabs: null 
            makecell: null 
            hyperref: null

        includes:
          in_header: extra_header.tex

bibliography: references.bib
fontsize: 12pt 
geometry: margin=2.5cm

---

@@

\newpage
# Introduction

A growing body of literature in economics explores how people develop incorrect beliefs 
about fundamentals. Most of this research centers on scenarios where agents 
passively observe the world, and incorrectly integrate the received information into their 
beliefs.\footnote{See \citet{benjamin2019} for a review of the literature on errors of 
probabilistic thinking}
However, many real-world situations cast agents as active participants in the generation of 
information. In these cases, the information they acquire is influenced by their actions and 
subsequent behavior is in turn determined by how they incorporate the information 
into their beliefs.

Consider a student who needs to decide how much effort to put into 
studying for an exam. 
Their decision will depend on two factors: their belief about their intrinsic 
ability, and their belief about how difficult the exam will be. The outcome 
they observe will be affected by how much they decide to study. Imagine that the student 
puts in a moderate amout of effort and gets a surprisingly good grade. Did they get a good grade because 
they are smarter than they thought? Or because the exam was easier than they had anticipated? 
Their future exam preparation strategy will depend fundamentally on which line of reasoning they 
take. This feedback loop is refered to as an \emph{endogenous information 
process}.

I compare a set of theories that model learning in an endogenous information setting and 
which can rationalize 
the persistence of misspecified beliefs. I describe a unifying framework that nests multiple 
theories of model learning and generates testable predictions for each of them. 
I cast this framework in a laboratory experiment design and test these predictions to identify which 
of these theories are consistent with the behavior I observe in the lab.
The framework features an agent who needs to 
learn two parameters: one that regards their own characteristics 
(an \emph{ego-relevant} parameter), 
and an exogenously determined state. In the context of the student preparing for exams, 
the two parameters  would be their intrinsic ability and the difficulty of the exam.

Misspecified beliefs about an ego-relevant parameter are often referred to as overconfidence 
(or underconfidence) and 
have been documented by behavioral scientists\footnote{\citet{kelley1980} 
provides a review of the psychology literature.} and economists\footnote{See \citet{benjamin2019} 
for a review of the literature in economics.} in a variety of settings.  @Oster2013 find that 
subjects who are at high risk of having Huntington's disease overestimate their probability of 
being healthy and make retirement decisions as if they were healthy. @Hoffman2020 show that workers 
overestimate the quality of their match to their 
current employment and are unlikely to look for other opportunities. @Camerer1999 find that 
entrepreneurs are overconfident about the quality of their enterprise, 
which leads to excessive entry and early exit from markets.

In all of these examples,
holding an incorrect belief about a fundamental leads to suboptimal choices with 
potentially high costs. In spite of the abundance of evidence, the scope of the existing research 
is limited in terms of the frameworks it considers. Most of the experimental evidence 
documenting the bias is collected in settings where subject are passive learners.\footnote{
\citet{Gotte2022} and \citet{Ozyilmaz2022} are exceptions that study settings with endogenous information processes.} They 
observe a noisy signal and report their beliefs over subsequent rounds.\footnote{\citet{Bracha2012} and 
\citet{Mobius2022} are some examples}. Although these studies provide important insights, they are 
not flexible enough 
to incorporate the richer theories that have been proposed more recently. In particular, they 
do not allow the studies of endogenous learning or for learning about multiple parameters at once.\footnote{\citet{Coutts2020} 
studies an environment with an ego-relevant parameter and an exogenous state but does not 
incorporate the endogenous information process.}

In my experiment, I move away from the standard framework to study a richer set of 
underlying mechanisms. In particular, 
the interaction between the two parameters, together with an endogenous information process, 
give rise to three possible explanations
for the persistence of incorrect beliefs: dogmatism, the existence of learning traps and misattribution bias.

Within the first explanation, even with perfect Bayesian updating the agent 
may fall into learning traps.\footnote{Conditions under which this happens are detailed in \citet{Hestermann2021}} 
These traps are characterized by a 
combination of misspecified beliefs and 
optimal actions which produce information that confirms the misspecified belief---this is a 
\emph{self-defeating equilibrium}. 
Once an agent falls into it, the belief will be stable and even with 
a correctly specified model of learning, they will not be able to abandon their misspecified beliefs. 
If, in addition the agent is 
dogmatic about their belief, @Heidhues2018 show that they will inevitably fall into a trap
and thus will be able to rationalize and sustein their initial dogmatically overconfident 
(or underconfident) belief.\footnote{\citet{Gotte2022} study the case of agents with dogmatic initial beliefs in a 
laboratory experiment.}

@Ba2023 moves away from dogmatism and endows the agents with a mechanism through which 
they can abandon misspecified beliefs; this allows them to avoid falling into learning traps. 
To do so agents perform Bayesian hypothesis tests to evaluate which is the more 
likely parameter out of two possibilities. By doing this, she characterizes the set of situations 
in which, even agents who consider alternative paradigms, may become trapped.\footnote{A 
similar mechanism is proposed by \citet{Schwarstein2021} in a setting with persuasion.}

Lastly, misattribution bias is the more classical explanation which has been widely studied in 
behavioral science.\footnote{see \citet{kelley1980} for a review.}
Agents who suffer from misattribution bias will attribute successes to their own ability---the 
ego-relevant parameter---and failures to bad luck---the state. 
Under this model of learning, even an agent who initially has a correct initial 
belief may become overconfident if they observe a sequence of successes. In this case, 
the main driver of the bias is not an initial misspecification or the presence of learning traps, 
it is the updating procedure itself.\footnote{A more general framework that can be used to model 
this bias has also been proposed by \citet{Brunnermeier2005} and empirically studied by \citet{Bracha2012}.}

These theories provide the main building blocks for a simplified framework that can be directly 
implemented in a laboratory experiment. The goal is to 
identify the which of the 3 forces---the presence of learning traps, mispecified initial beliefs, or missatribution 
bias---better explains the observed behavior. To determine the fit of the models, I compare the behavior
predicted by each theory to the benchmark given by the fully Bayesian updating procedure.

I also study whether the misspecification is inherently linked to the ego-relevance the parameters 
or a more general phenomenon. I vary the degree of ego-relevance by introducing a 
treatment in which subjects learn about the ability of another participant.
In this treatment, the participants know only the gender and nationality of the other, which can 
induce stereotypes---a different type of misspecification.  
If correct learning about the parameters happens at higher rates in the stereotype treatment, 
it would suggest that the 
bias is intrinsically linked to the ego-relevance of the parameter. In contrast, if similar 
biased behavior arises in both treatments, it is more likely that the main driver of these types 
of misspecified beliefs is the updating procedure itself or the endogenous information process.

Although some agents do fall into learning traps, I find that the behavior of most 
subjects is better explained by misattribution bias. I also find that misattribution is 
no more prevalent in the ego-relevant than in the stereotype condition. This suggests 
that the main driver of the misspecification is the updating proceduce. However, the underlying 
mechanism by which the bias is generated may be 
different in both treatments: While in the ego-relevant condition, the intuitive explanation behind the
bias is that the agent prefers to hold themselves in high esteem, in the stereotype condition
it seems to be driven by some sort of self-censoring---when subjects realize that they underestimated 
the ability of another participant based on their gender and nationality, they over-compensate by
overestimating their ability.

Finally, I estimate the structural parameters of the models to study model-heterogeneity in the sample.
I find that even at an individual level the behavior is better explained by misattribution bias for most
subjects. There is a smaller group of subjects that can be better explained through dogmatic beliefs
and hypothesis testing and none of them behave in line with the fully Bayesian benchmark.

In what follows I first discuss the theoretical framework and the predictions of each of the theories.
Then I introduce a unifying example and my hypotheses. I then describe the experimental
design and the results, as well as the estimation of the parameters.

@@

The theories that I consider make predictions in two distinct frameworks. I first detail each theory 
within their original framework and then develop a unifying example that allows me to
compare the predictions of all of them. In framework 1 I consider a setting in which the an agent observes
a continuous output and infers the underlying state. In framework 2, the agent observes a binary 
outcome and infers the underlying probability of success. In both frameworks, I consider two nested theories 
of learning.

@@

_The student must choose an effort level $a$, which determines her grade. For whatever the chosen effort level, 
the student must pay a cost $c(a) = \frac{1}{2}a^2$ and observes a grade of $(\theta^*+a)\omega^*$. She repeats 
this process for infinitely many periods. 
Assume also that the student's prior is such that $\mathbb{E}[\omega]= \omega^*$ and she is dogmatic 
about being of 
type $\hat{\theta}>\theta^*$.\footnote{The example is illustrated for an overconfident student but 
the results are symmetric for a dogmatic student who initially 
places probability one on some $\tilde{\theta}<\theta^*$.}
Therefore, the student's payoff in period $t$ is given by _

@@

_Under this specification, the myopic optimal effort level is $a_t^* = \omega^*$. 
Because the student does not know $\omega^*$,
he will choose $a_t = \mathbb{E}_t(\omega)$ where the expectation is taken with respect to the 
student's belief at the beginning of
period $t$. If she does not revise his effort choice for $k$ periods, she will receive 
an average utility of 
$(\theta^*+a_t^*)\omega^* - \frac{1}{2}a_t^{*2}$ but he was expecting an average utility of 
$(\hat{\theta}+a_t^*)\omega^* - \frac{1}{2}a_t^{*2}$. In response, she will apply Bayes 
rule to update his beliefs about
$\omega$ to get the posterior belief with 
$\mathbb{E}_{t+k}[\omega] = \frac{(\theta^{*} + \omega^{*})\omega^{*}}{\hat{\theta} + \omega^{*}}$
which is lower than the initial belief. This will cause the student to choose a lower effort at $t+k$. As a result, she will again
receive an average utility that is lower than what he expected which will cause her belief to drift further down. This process will
continue until the average utility equals her expected utility under the dogmatic belief that assigns probability 1 to $\hat{\theta}$. 
At that point, the student will have reached a self-defeating equilibrium and she will continue to choose sub-optimal effort forever. _

@@

The overall pattern is that bad news are associated with a decrease in effort, and a 
more negative news differential is associated with a larger decrease in effort. On the contrary, good news are associated with 
a small decrease in effort. As mentioned in the predictions section, this is consistent with the self-attribution bias as well
as with a change of paradigm. It is not consistent with the dogmatic modeler or the Bayesian. Because 
I do not directly elicit beliefs, it is not possible to distinguish between the self-attribution bias and the change of 
paradigm with this data alone.
Nevertheless, the fact that about a third of the subjects started with a correctly specified belief and ended up overconfident
is consistent only with the self-attribution bias. this points towards the fact that 
misattribution bias is the most likely explanation for the behavior of the subjects in the experiment.

@@

A regression analysis of this same data reveals that indeed, the slopes are different for godd and bad
news and that the difference is statistically significant.\footnote{I run a robustness check where I define
good news to be a net-positive sample (half or more of the outcomes observed in the sample received in that round
were successful) and bad news to be net-negative sample. I find the same pattern in this alternative regression model
and the results are reported in the appendix.} 
The regression results are reported in Table \ref{tab:regression}.
columns 2 and 3 of the table consider the model for each treatment separately. I find no significant differences in 
the parameters across treatment. This is consistent with the fact that the bias is not driven by the ego-relevance 
of $\theta$.\footnote{The ego-relevance of $\theta$ is the motivation behind the self-attribution bias, 
however, the bias itself is due to an incorrect updating procedure in response to good and bad news. The presence of 
asymmetric updating in the stereotype condition can be due to self-censoring of the subjects. In particular, subjects 
who initially underestimated the performance of the other participant might have realized their mistake and 
by overestimating it after good news. This is consistent with the data presented here as well as with 
the data observed in the transition matrix for stereotypes which is reported in the Appendix}

@@

More convincing evidence against the Switching model is the lack of inverse or attenuated correlation between 
the change in effort and the news differential. In particular, if the switcher were the underlying model,
I would expect overconfident subjects who start with an overestimation of $\theta$ to increase their effort
in response to surprising bad news. This would flatten the slope of the regression line 
for bad news relative to what a Bayesian would do. \ref{tab:regression} shows that the slope for bad news for
the simulation\footnote{The details on how the data was simulated are explained in the Appendix} 
of the Bayesian updates is much flatter than the slope in the data.

@@

The switching threshold is the parameter that determines the likelihood ratio at which the switcher
abandons the status quo paradigm and adopts the alternative. In order to estimate this parameter, I
assume that all subjects are switchers and look at the rounds in which they chose to reveal a probability
matrix that was different from the one they revealed in the previous round. I then calculate the likelihood
ratio for the alternative paradigm to the status quo. I average across all such 
likelihood ratios and use the average as an estimate of the switching threshold.

In detail, Let $b_{t}^\theta$ be the belief about $\theta$ in period $t$.\footnote{in the experimental data, 
I use the revealed matrix as a proxy for the belief. I take the matrix revealed in the previous period 
to be the status-quo and the matrix revealed in period $t$ to be the alternative}. 
And consider a subject for whom $b_{\tau-1}^\theta \neq b_{\tau}^\theta$, 
where $b_{t}^\theta$. That is, in period $\tau-1$ they chose to reveal a matrix that was different from the one they
revealed in period $\tau$. The matrix revealed in period $\tau-1$ is the status quo paradigm and the matrix revealed in
period $\tau$ is the alternative paradigm. The likelihood ratio for the alternative paradigm to the status quo is
$$
\ell = \frac{L(\theta | b_{\tau}^\theta)}{L(\theta | b_{\tau-1}^\theta)}=\frac{p_\tau(o_t|b_{\tau}^\theta)}{p_{\tau}(o_t|b_{\tau-1}^\theta)} 
$$

where $p_\tau(o_t|b_{\tau}^\theta)$ is the probability of observing the signal $o_t$ when the agent is of type $b_{\tau}^\theta$.

The estimate of the switching threshold, $\alpha$, is the average of the likelihood ratios across all subjects who switched
in the experiment, which is given by

$$\hat{\alpha} = \frac{1}{N}\sum_{i=1}^N \ell_{i\tau}$$

Where $N$ is the number of switches observed in the experiment.

Because the feedback that subjects receive is the equivalent of 10 signal realizations, the estimate 
of the parameter is an upper bound of the true parameter. A lower bound would be given by the
average likelihood ratio in the round before the switch happens. I choose to estimate the upper bound
to impose a harsher test on the theory.

Because the likelihood ratio is a function of the signal, it depends on the sequence of signals that 
subjects observe. This means that in some cases, the likelihood ratio increases by a large amount from 
one period to another. Although such cases are not common\footnote{In the data, only 1.8% of the 
observations have a likelihood ratio that exceeds 5.}, a few very large values can skew the average. 
To avoid a biased parameter, I restrict my sample to those observations for which the likelihood ratio
is less than 5.\footnote{The full distribution of likelihood ratios is reported in the appendix}
Under these assumptions, the estimate of the switching threshold is `r round(mean(updates[lr < 5 & change_indicator==1, lr]), 2)`.

This means that the evidence supporting the alternative paradigm needs to be approximately 10% more
compelling than the evidence supporting the status quo paradigm in order for the agent to switch.
The estimation treatment by treatment is slightly larger for the Ego condition than for the Stereotype
condition. However, the difference is not statistically significant (p-value=0.96).

I then use this estimate to simulate the behavior of the switcher to assess the fit of the model.
Which is discussed in the next section.

@@

Estimation of the parameters of the self-attribution bias is more involved than estimation of the switching
threshold. In this case, there is no obvious identification strategy that provides a simple estimate 
of the parameters. Instead, I use the simulated method of moments to estimate the parameters of the model.

In particular, I simulate the behavior of the self-attribution bias for each possible prior and over 
a grid of parameters. I make the simplifying assumption that the distortion is symmetric across good and bad news.
That is, I assume that $c(\theta_H, \omega, \text{good news})=c(\theta, \omega_L, \text{bad news})=:c_g$ and that 
$c(\theta_M, \omega, \text{bad news})=c(\theta, \omega_M, \text{good news})=:c_b$. In addition, I normalize the 
remaining cases to be equal to 1. This leaves me with only 2 parameters to estimate.

The moments that I match in the data are the reaction to good news and the reaction to bad news. In particular,
I match the average change in effort in response to good news and the average change in effort in response to bad news.
That is I match the following moments:

$$
\bar{G} = \frac{1}{N}\sum_{i=1}^N \frac{1}{T}\sum_{t=1}^T \mathbb{I}\{o_t > \bar{o}\} \cdot (e_t - e_{t-1}) 
$$
$$
\bar{B} = \frac{1}{N}\sum_{i=1}^N \frac{1}{T}\sum_{t=1}^T \mathbb{I}\{o_t < \bar{o}\} \cdot (e_t - e_{t-1})
$$
where $N$ is the number of subjects in the sample, $T$ is the number of rounds in the experiment.

I then use the simulated method of moments to estimate the parameters of the model. In particular, I
simulate the behavior of the model for each possible prior and over a grid of parameters. I then
calculate the average change in effort in response to good news and the average change in effort in
response to bad news for each possible prior and over the grid of parameters and across $I$ simulated 
indiviaduals. 
I then minimize the objective function given by the sum of squared differences between the simulated
moments and the empirical moments. That is, I solve the following problem

# Model fit

@@

I have small or no effects throughout on the ego-relevance of the type. Although the motivation behind the 
theories does not directly apply to the case of stereotypes, the lack of stark differences suggests that 
biases arise due to an incorrect updating procedure and not necessarily due to the self-serving motivation 
behind the bias. The lack of differences in the behavior of the subjects across treatments could also 
be due to self-censoring of the subjects. In particular, subjects might be initially reluctant to 
express their true belief about a particular group of people. In addition, as they learn and realize 
that their beliefs might be incorrect, they might overreact to signals to try to compensate for the initial 
bias. Such an updating procedure would be consistent with the self-attribution bias for the case of stereotypes.
I remain agnostic about whether that is truly the case or whether the lack of differences is due to other 
factors.

@@

\newpage
\appendix

# appendix
## Transition matrix by treatment

@@

## Full distribution of likelihood ratios
The likelihood ratios at rounds in which I observe a change in the belief about $\theta$ presents a
long right-tail. In particular, 1.8% of the observations have a likelihood ratio that exceeds 5 
which would mean that the agent requires evidence that is 5 times more compelling in favor 
of the alternative paradigm in order to switch. Figure \ref{fig:lr-hist} shows the distribution of
likelihood ratios for all observations.

![Distribution of likelihood ratios for all observed belief changes](../figures/lr_hist.png){#fig:lr-hist}

@@

## Reaction to good and bad news (robustness check)

I use an alternate definition for good and bad news and check whether the results are robust to this. 
I define a signal to be a net positive if the number of successes is equal to or greater than the number of failures.
I define a signal to be a net negative if the number of failures is greater than the number of successes.
I now treat net positive signals as good news and net negative signals as bad news. I then rerun the regression 
analysis and find that the results are robust to this alternative definition. In particular, net negative 
signals are associated with a stark decrease in effort for both treatments. Net positive signals are
associated with a small decrease in effort for both treatments. The results are reported in Table \ref{tab:regression-net}.

```{r, echo=FALSE, message=FALSE, warning=FALSE }
reg2 <- lm(effort_change ~ positive_signal + signal_prev + signal_prev:positive_signal, data=updates[round_number>1,])
robust_se <- sqrt(diag(vcovHC(reg)))

# Attach the robust standard errors to the coefficient estimates
coefs2 <- coef(reg2)
coefs2 <- cbind(coefs2, robust_se)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# run the regression only for treatment  = ego
reg2_ego <- lm(effort_change ~ positive_signal + signal_prev + signal_prev:positive_signal, data=updates[round_number>1 & treatment == 'ego',])

# Attach the robust standard errors to the coefficient estimates
coefs2_ego <- coef(reg2_ego)
coefs2_ego <- cbind(coefs2_ego, robust_se)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# run the regression only for treatment  = stereotype
reg2_stereotype <- lm(effort_change ~ positive_signal + signal_prev + signal_prev:positive_signal, data=updates[round_number>1 & treatment == 'stereotype',])

# Attach the robust standard errors to the coefficient estimates
coefs2_stereotype <- coef(reg2_stereotype)
coefs2_stereotype <- cbind(coefs2_stereotype, robust_se)
```

stargazer(reg2, reg2_ego, reg2_stereotype,  
          title='Regression of effort change on news difference with robust standard errors', 
          dep.var.labels='Change in effort', 
          covariate.labels=c('Net-positive signal', 'Signal value', 'Signal value * Net-Positive signal'),
          column.labels=c('All', 'Ego-relevant', 'Stereotype'), 
          omit.stat=c('f', 'ser'), 
          digits=2, 
          label='tab:regression-net', 
          header=FALSE, 
          se = list(coefs2[, 2], coefs2_ego[, 2], coefs2_stereotype[, 2]))

# Experimental Instructions

@@

_The full data generating process is printed on the rear side of the instructions and always available to the subjects throughout part 2._
_Each of the matrices is labeled with "if your/the other participant's score is..." depending on the treatment_

@@

\newpage
# References
